\documentclass[12pt]{article}

\usepackage{helpers}

\begin{document}

Name: \makebox[3in]{\hrulefill
%NAME
\hrulefill}

\vfill

\begin{center}
{\huge Test 4}

{\large 2026}
\end{center}

\testrules

\vfill

\pledge

\newpage

\section*{Memory Hierarchy}

Demonstrate understanding of caches and the memory hierarchy, through answering conceptual questions and describing processes.\\


\begin{enumerate}

\item What does latency mean in the context of memory?
\vfill

\item List the six types of storage we've covered in \textbf{decreasing} order of latency.
\vfill

\item What does ``random access'' mean in Random Access Memory?
\vfill



\newpage
\item Explain the difference between a HDD and an SSD.
\vfill

\item What are temporal locality and spatial locality, and how are they used by the CPU cache to run programs more efficiently?
\vfill

\item What is a cache hit? A cache miss? A cache eviction?
\vfill




\end{enumerate}

\standardsfooter

\newpage

\section*{Operating Systems}

Demonstrate understanding of the operating system and multicore programming, through answering conceptual questions and describing processes.

\begin{enumerate}
\item What is the role of the operating system in a computer?
\vfill

\item How does context switching work?
\vfill

\item What does it mean for a process to be a parent of another process?
\vfill

\newpage

\item How is a GPU different from a CPU?
\vfill

\item Explain the difference between multiprogramming and multicore programming.
\vfill

\item Explain the differences between processes, threads, and cores.
\vfill

\end{enumerate}

\standardsfooter

\end{document}