\documentclass[12pt]{article}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\parindent}{0in}
\setlength{\parskip}{\baselineskip}

\usepackage{amsmath,amsfonts,amssymb,graphicx,xcolor,mathtools,mathabx,siunitx}
\sisetup{tight-spacing=true}

\newcommand{\purple}[1]{{\color{purple} #1}}

\begin{document}

PHYS 374 Fall 2020\hfill Worksheet 10: Lagrangian Mechanics in an Electromagnetic Field\\
\\
Name:\\
\\
Please submit as a PDF on Moodle. Include any calculations made using external tools.

\hrulefill
\\
\\
A particle with mass $m$ and charge $q$ begins at rest in a constant electric field $\vec{E}$ and a constant magnetic field $\vec{B}$. The electric field points in the $x$ direction and the magnetic field points in the $y$ direction. 
\begin{enumerate}
    \item What is the vector potential $\vec{A}$ for this system? There are several possible answers.
    \item What is the scalar potential $V$ for this system?
    \item What is the Lagrangian for this system? 
    \item Show that the equations of motion are as follows:
    \begin{align*}
        m \ddot{x} &= qE - qB\dot{z} &
        m \ddot{y} &= 0 &
        m \ddot{z} &= qB \dot{x}
    \end{align*}
    \item Perform a change of variables from $x$ and $z$ to $v_x=\dot{x}$ and $v_z=\dot{z}$. Then show that the equations of motion can be rewritten in the familiar forms below:
    \begin{align*}
        \ddot{v_x} &= -\omega^2 v_x &
        \ddot{v_z} &= -\omega^2 v_z + \alpha
    \end{align*}
    \item Solve the equations of motion for $v_x(t)$ and $v_z(t)$, and use initial conditions to find the value of any undetermined constants. Confirm that your solutions satisfy the original (coupled) equations of motion. 
    \item What does the motion look like? Draw a picture. 
\end{enumerate}

For reference:
\begin{align*}
    \mathcal{L} &= \tfrac{1}{2} m \dot{\vec{r}}^2 - qV + q \dot{\vec{r}} \cdot \vec{A} &
    \vec{B} &= \nabla \times \vec{A} &
    \vec{E} &= -\nabla V - \frac{\partial \vec{A}}{\partial t} &
\end{align*}
And in rectangular coordinates:
$$
\nabla \times \vec{A} = 
\left[
{\begin{array}{c}
    \partial_x \\
    \partial_y \\
    \partial_z
  \end{array} }
\right]
\times
\left[
{\begin{array}{c}
    A_x \\
    A_y \\
    A_z
  \end{array} }
\right]
=
\left[
{\begin{array}{c}
    \partial_y A_z - \partial_z A_y \\
    \partial_z A_x - \partial_x A_z \\
    \partial_x A_y - \partial_y A_x
\end{array} }
\right]
\quad\quad\text{where $\partial_x=\frac{\partial}{\partial x}$}
$$

\end{document}