\documentclass[12pt]{article}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\parindent}{0in}
\setlength{\parskip}{\baselineskip}

\usepackage{amsmath,amsfonts,amssymb,graphicx,xcolor,mathtools}

\newcommand{\purple}[1]{{\color{purple} #1}}

\begin{document}

PHYS 374 Fall 2020\hfill Worksheet 6: Hanging Chain\\
\\
Name:\\
\\
Please submit as a PDF on Moodle. Include any calculations made using external tools.

\hrulefill
\\
\\
Two rigid posts of height $H$ stand at $x=-L$ and $x=L$. A chain of length $\ell$ hangs between them. The mass of the chain, $m$, is uniformly distributed along its length.
\begin{enumerate}
\item Draw a picture!
\item Write down the potential energy of a small section of the chain $dm$, then write down an integral over $dm$ for the potential energy of the entire chain.
\item Rearrange your potential energy integral into rectangular coordinates. 
\item The length of the chain is fixed. Write down that constraint as another integral, then rearrange it into the same coordinates as above.
\item Combine your two integrals into a single expression using a Lagrange multiplier $\lambda$.
\item Use the Euler-Lagrange equation to find a differential equation describing the chain.

Hint: recall when $f(y, y', x)$ doesn't depend on $x$, the EL equation can be rewritten: 
$$
\frac{\partial f}{\partial y} - \frac{d}{dx} \frac{\partial f}{\partial y'} = 0
\quad\quad\rightarrow\quad\quad
f - y' \frac{\partial f}{\partial y'} = \text{constant}
$$
\item Solve the differential equation to get an expression $y(x)$.

Hint: try the substitution $\cosh u = \alpha y + \beta$, where $\alpha$ and $\beta$ are constants. Note also:
\begin{align*}
    \tfrac{d}{du} \cosh u &= \sinh u&
    \tfrac{d}{du} \sinh u &= \cosh u&
    \cosh^2 u - \sinh^2 u &= 1
\end{align*}
\item Explain the significance of any undetermined constants in your expression $y(x)$. Give a simple expression relating each constant to known quantities (you do not need to solve those expressions). 
\end{enumerate}
\end{document}